Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-11-26 WSSSPE3/report/wg_appendix_software_credit.tex
small edits to edits

During the first day of WSSSPE3 breakout sessions, the group discussed the
Entertainment Identifier Registry (EIDR)~\cite{EIDR} as a potential model for
scientific software. That system assigns unique Digital Object Identifiers
(DOIs)---the same system used for scientific publications---to all content
(e.g., movies, television shows) and contributors, along with relevant metadata.
One important use of this system is to track rights and credits for contributors
to entertainment works in order to distribute revenues---similar to the proposed
transitive credit concept.

@@

Finally, although a discrete task from software citations themselves,
significant discussion focused on ensuring software citations are indexed in the
same manner as publications, allowing the construction of a corresponding
software citation network. Currently, software releases can receive citable DOIs
via Zenodo~\cite{zenodo-web} and figshare~\cite{figshare-web}; however, these
citations are not processed by indexers such as Web of Science, Scopus, or
Google Scholar. Thus, either in parallel or following the primary task, the
group will need to reach out to these organizations. Initial conversations with
Elsevier\slash Scopus via Michael Taylor during WSSSPE3 clarified that Scopus
is not yet DataCite DOI aware, and also does not yet have an internal identifier
for software or data (but needs\slash plans to add this support). Taylor said
they prefer a ``software article'' with the usual article metadata (e.g.,
authors, citations), and mentioned Zenodo as an example -- this proposal seemed
to align with our group discussions. Taylor also mentioned another benefit of the
software and associated DOI on GitHub: in addition to a citation, one could obtain
statistics on usage/downloads/forks, which happens to be what
Depsy\footnote{Depsy: \url{https://depsy.org}} is beginning to try to do.

@@

Modern scientific research relies considerably on software, and ensuring that
developers of such scientific software receive credit for their efforts will
encourage additional creation and maintenance. \katznote{the prev sentence is
good, but feels a bit odd in the middle of the appendix, rather than at the
beginning} There currently is no standard mechanism for citing software or
receiving credit for software (akin to publications/citations). Software is
eligible for DOI assignment, but DOI metadata fields are not well tuned or
standardized for software (vs.\ publications). Some software providers apply for
DOIs, but the identifiers are still not widely adopted. Also, there is no mechanism to cite
software dependencies within software.

>>>>

During the first day of WSSSPE3 breakout sessions, the group discussed the
Entertainment Identifier Registry (EIDR)~\cite{EIDR} as a potential model for
scientific software. That system assigns unique Digital Object Identifiers
(DOIs)---the same system used for scientific publications---to all content
(e.g., movies, television shows) and contributors, along with relevant metadata.
One important use of this system is to track rights and credit for contributors
to entertainment works in order to distribute revenues---similar to the proposed
transitive credit concept.

@@

Finally, although a discrete task from software citations themselves,
significant discussion focused on ensuring software citations are indexed in the
same manner as publications, allowing the construction of a corresponding
software citation network. Currently, software releases can receive citable DOIs
via Zenodo~\cite{zenodo-web} and figshare~\cite{figshare-web}; however, these
citations are not processed by indexers such as Web of Science, Scopus, or
Google Scholar. Thus, either in parallel or following the primary task, the
group will need to reach out to these organizations. Initial conversations with
Elsevier\slash Scopus via Michael Taylor during WSSSPE3 clarified that Scopus
is not yet DataCite DOI aware, and also does not yet have an internal identifier
for software or data (but needs\slash plans to add this support). Taylor said
they prefer a ``software article'' with the usual article metadata (e.g.,
authors, citations), and mentioned Zenodo as an example -- this proposal seemed
to align with our group's discussions. Taylor also mentioned another benefit of the
software and associated DOI on GitHub: in addition to a citation, one could obtain
statistics on usage/downloads/forks, which happens to be what
Depsy\footnote{Depsy: \url{https://depsy.org}} is beginning to try to do.

@@

Modern scientific research relies considerably on software, and ensuring that
developers of such scientific software receive credit for their efforts will
encourage additional creation and maintenance. \katznote{the prev sentence is
good, but feels a bit odd in the middle of the appendix, rather than at the
beginning} There currently is no standard mechanism for citing software or
receiving credit for software (akin to publications/citations). Software is
eligible for DOI assignment, but DOI metadata fields are not well tuned or
standardized for software (vs.\ publications). Some software providers apply for
DOIs, but this is not widely adopted. Also, there is no mechanism to cite
software dependencies within software.
